---
title: "Getting started with 'ggmapinset'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with 'ggmapinset'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggmapinset)
library(ggplot2)
library(sf)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
```

This article provides some recipes for working with insets.

## Different aesthetics for inset layer
By default, `geom_sf_inset()` creates two copies of the map layer: one for the
base map and the other for the inset map. The inset is transformed and clipped,
but uses the same underlying aesthetics mapping and parameters.

If you want to have different aesthetics for the two layers, you'll need to turn
off this copying with `map_base = "none"`. With this parameter set and an
`inset` parameter provided, only the inset layer will be drawn. To draw only the
base layer, you can use `map_inset = "none"`, `inset = NULL`, or simply use the
normal `geom_sf()`.

```{r separate, fig.width=7, fig.height=3.5}
cfg <- configure_inset(centre = st_centroid(st_geometry(nc)[nc$NAME == "Bladen"]),
                       scale = 1.5, translation = c(-180, -50), radius = 50, units = "mi")

ggplot(nc) +
  # this is equivalent to the following line:
  # geom_sf_inset(fill = "white") +
  geom_sf(fill = "white") +
  geom_sf_inset(aes(fill = AREA), inset = cfg, map_base = "none") +
  geom_inset_frame(inset = cfg) +
  coord_sf()
```

## Inset frame backgrounds
If you'd like to add a background to the inset frame (perhaps because it
overlaps with part of the map), this can be done with the normal aesthetics.
The frame is composed of the two circles then the two lines connecting them, so
we want `fill = c(NA, "white", NA, NA)` to fill just the large circle.
Note that under some circumstances the frame data will have different
dimensions, due to facets (the frame is repeated for each facet) or when the
two circles are nested (in which case the lines are not drawn).

We want the filled frame to sit in between the base map and the inset layers.
```{r frame_fill, fig.width=7, fig.height=3}
cfg <- configure_inset(centre = st_centroid(st_geometry(nc)[nc$NAME == "Yancey"]),
                       scale = 2, translation = c(100, -120), radius = 50, units = "mi")

ggplot(nc) +
  geom_sf(aes(fill = AREA)) +
  geom_inset_frame(inset = cfg, fill = c(NA, "white", NA, NA)) +
  geom_sf_inset(aes(fill = AREA), inset = cfg, map_base = "none") +
  coord_sf()
```

