---
title: "Getting started with 'ggmapinset'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with 'ggmapinset'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggmapinset)
library(ggplot2)

nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
```

This article provides some recipes for working with insets.

## Different aesthetics for inset layer
By default, `geom_sf_inset()` creates two copies of the map layer: one for the
base map and the other for the inset map. The inset is transformed and clipped,
but uses the same underlying aesthetics mapping and parameters.

If you want to have different aesthetics for the two layers, you'll need to turn
off this copying with `map_base = "none"`. With this parameter set and an
`inset` parameter provided, only the inset layer will be drawn. To draw only the
base layer, you can use `map_inset = "none"`, `inset = NULL`, or simply use the
normal `geom_sf()`.

```{r separate, fig.width=7, fig.height=3.5}
ggplot(nc) +
  # this is equivalent to the following line:
  # geom_sf_inset(fill = "white", map_inset = "none") +
  geom_sf(fill = "white") +
  geom_sf_inset(aes(fill = AREA), map_base = "none") +
  geom_inset_frame() +
  coord_sf_inset(inset = configure_inset(
    centre = sf::st_centroid(sf::st_geometry(nc)[nc$NAME == "Bladen"]), scale = 1.5,
    translation = c(-180, -50), radius = 50, units = "mi"))
```

## Inset frame backgrounds
By default, the inset frame is transparent, although often it makes sense to add
a solid background so that the inset is distinguishable from any overlapping part
of the base map. The aesthetics of the two parts of the frame and the burst lines
connecting them can be controlled separately.

Note that when the background is filled, we need to specify the base and inset
maps in separate layers so that the frame can slip in between them.
```{r frame_fill, fig.width=7, fig.height=3}
ggplot(nc) +
  geom_sf(aes(fill = AREA)) +
  geom_inset_frame(target.aes = list(fill = "white")) +
  geom_sf_inset(aes(fill = AREA), map_base = "none") +
  coord_sf_inset(inset = configure_inset(
    centre = st_centroid(st_geometry(nc)[nc$NAME == "Yancey"]), scale = 2,
    translation = c(100, -120), radius = 50, units = "mi"))
```

## Multiple insets
For multiple insets, the appropriate inset configuration just needs to be passed to each
layer separately. It's probably clearer to avoid providing an inset to the coordinate
system in this case.

Since the inset-aware layers will duplicate themselves for the base and inset maps,
you will probably want to disable that behaviour with `map_base = "none"` to avoid
having multiple identical copies of the base map.
```{r multiple, fig.width=7, fig.height=5}
inset1 <- configure_inset(
  centre = sf::st_centroid(sf::st_geometry(nc)[nc$NAME == "Bladen"]), scale = 1.5,
  translation = c(150, -50), radius = 50, units = "mi")
inset2 <- configure_inset(
  centre = sf::st_centroid(sf::st_geometry(nc)[nc$NAME == "Orange"]), scale = 3,
  translation = c(30, 120), radius = 30, units = "mi")

ggplot(nc) +
  # base map
  geom_sf_inset() +
  # inset 1
  geom_sf_inset(map_base = "none", inset = inset1) +
  geom_inset_frame(inset = inset1, colour = "red") +
  # inset 2
  geom_sf_inset(map_base = "none", inset = inset2) +
  geom_inset_frame(inset = inset2, colour = "blue")
```

